
Pro memorie :

meeting 2025-04-08  : https://drive.google.com/drive/folders/1J16lJqVCQ_Txaiix2DBOPZi1mdyYOQ4B
C2 log  : https://drive.google.com/drive/folders/1SwqbaRRxpQv4ByoCZskPje64NW5Es-vE?usp=sharing

Exact KPI formulation (at the end of the project):

End-of-project
KPI = Reduction of population size in managed sites of emerging plant species, from start of intervention (as baseline) to 2026.
Target = ≥ 80% decrease on baseline

5-years-after-project
KPI = A final assessment will be made at the end of the AfterLife project (2031) to check for complete population eradication.
Target = 100% decrease on baseline

Since a project baseline needs to be given as an absolute number, the choice is made to express these as number of m². This is taken as the sum of currently known populations under C2 management. For the populations currently expressed in number of plants (or other), a conversion factor will be applied (cover of a single plant).

# PREPARATION

```{r prep packages}
library(tidyverse)
library(readxl)
library(stringr)
```

```{r prep memory}
rm(list = ls())
```

# DATA

```{r read data}
file <- "./data/input/C2 RapidResponseActionsLog_asof26-08-04.xlsx" # most recent version

raw.pop <- read_excel(file, sheet = 2, skip = 1)
raw.act <- read_excel(file, sheet = 3, skip = 1)
```

```{r select}
pop <- raw.pop %>% 
  filter(group == "plants") |> 
  transmute(
    popcode,
    species = str_extract(species, "^[^ ]+"), # genus
    # latitude,
    # longitude,
    # region,
    before_m2 = `INITIAL_infested_m²`,
    before_other = INITIAL_size_quantity,
    before_other_unit = INITIAL_size_unit,
    after = REMAINING_size,
    after_unit = REMAINING_size_unit,
    eradication_status
    )
```

```{r}
spp <- unique(pop$species)
```

```{r fix and complete}

# Eg4, Eg5 : assume "<1" ~ 1
pop$before_m2[pop$before_m2 == "<1"] <- "1"

# Eg3 : data missing, let's assume 3m² (based on wnm.be and field visits)
pop$before_m2[pop$popcode == "Eg3"] <- "3"

# other
pop$after[pop$after == "n/a"] <- NA

pop$before_m2 <- as.numeric(pop$before_m2)
pop$after <- as.numeric(pop$after)

```

# CONVERT ALL TO M²

```{r}
La.plants.to.m2 <- 1.2 # Vibe
Lp.plants.to.m2 <- 0.7 # Vibe
Lg.m3.to.m2     <- 6   # Vibe
Ch.m3.to.m2     <- 35  # Vibe
Ed.m3.to.m2     <- 4.6 # Vibe
Eg.plants.to.m2 <- 0.2 # Vibe
Pj.plants.to.m2 <- 1.8 # Vibe
Kp.plants.to.m2 <- 1.0 # Vibe
```

## 'BEFORE' in M²

```{r}
pop.m2 <- pop %>%
  # Lysichiton
  mutate(before_m2 = ifelse(species == "Lysichiton" &
                              is.na(before_m2) &
                              before_other_unit == "plants",
                            before_other * La.plants.to.m2,
                            before_m2)) |> 
  # Ludwigia
  mutate(before_m2 = ifelse(species == "Ludwigia" &
                              is.na(before_m2) &
                              before_other_unit == "plants",
                            before_other * Lp.plants.to.m2,
                            before_m2)) |> 
  mutate(before_m2 = ifelse(species == "Ludwigia" &
                              is.na(before_m2) &
                              before_other_unit == "m3",
                            before_other * Lg.m3.to.m2,
                            before_m2)) |> 
  # Crassula
  mutate(before_m2 = ifelse(species == "Crassula" &
                              is.na(before_m2) &
                              before_other_unit == "m3",
                            before_other * Ch.m3.to.m2,
                            before_m2)) |> 
  # Egeria
  mutate(before_m2 = ifelse(species == "Egeria" &
                              is.na(before_m2) &
                              before_other_unit == "m3",
                            before_other * Ed.m3.to.m2,
                            before_m2))
```

## 'AFTER' in M²

```{r}
pop.m2 <- pop.m2 |> 
  mutate(after_m2 = NA, .before = after) |>
  # logic : if considered eradicated, then surface area = 0 m²
  mutate(after_m2 = ifelse(eradication_status == "ERADICATED",
                           0,
                           after_m2)) |>
  # Lysichiton
  mutate(after_m2 = ifelse(species == "Lysichiton" &
                             is.na(after_m2) &
                             after_unit == "plants",
                            after * La.plants.to.m2,
                            after_m2)) |> 
  # Ludwigia
  mutate(after_m2 = ifelse(species == "Ludwigia" &
                             is.na(after_m2) &
                             after_unit == "m3",
                            after * Lg.m3.to.m2,
                            after_m2)) |>
  # Crassula
  mutate(after_m2 = ifelse(species == "Crassula" &
                             is.na(after_m2) &
                             after_unit == "m3",
                            after * Ch.m3.to.m2,
                            after_m2)) |> 
  # Erythrante
  mutate(after_m2 = ifelse(species == "Erythranthe" &
                             is.na(after_m2) &
                             after_unit == "m2",
                           after,
                           after_m2)) |> 
  mutate(after_m2 = ifelse(species == "Erythranthe" &
                             is.na(after_m2) &
                             after_unit == "plants",
                            after * Eg.plants.to.m2,
                            after_m2)) |> 
  # Petasites
  mutate(after_m2 = ifelse(species == "Petasites" &
                             is.na(after_m2) &
                             after_unit == "plants",
                            after * Pj.plants.to.m2,
                            after_m2)) |> 
  # Koenigia
  mutate(after_m2 = ifelse(species == "Koenigia" &
                             is.na(after_m2) &
                             after_unit == "plants",
                            after * Kp.plants.to.m2,
                            after_m2))
```

```{r tdot}
missing <- pop.m2 |>
  filter(is.na(after_m2)) |>
  select(popcode) |>
  left_join(raw.pop |> select(popcode, region))

write.csv2(missing, "./data/missing_asof26-08-05_td.csv")
```

```{r strip it}
pop.strip <- pop.m2 |> 
  select(popcode,
         species,
         before_m2,
         after_m2) |> 
  filter(!is.na(after_m2)) |> 
  mutate(
    reduction = 1 - (after_m2 / before_m2)
    ) |> 
  # accounting for the few populations that seem to have grown
  mutate(
    after_m2.cap = ifelse(after_m2 > before_m2,
                          before_m2,
                          after_m2), .after = after_m2) |>
  mutate(
    reduction.corr = 1 - (after_m2.cap / before_m2)
    )
```

# KPI

```{r naive method}
# using the capped after_m2
kpi.1 <- 1 - sum(pop.strip$after_m2.cap) / sum(pop.strip$before_m2)
kpi.1
```

```{r per-population method}
kpi.2 <- mean(pop.strip$reduction.corr)
kpi.2
```

# OVERVIEW of ERADICATION STATUS

```{r}
# only proceed with the populations that have before/after data
pop.sel <- pop.strip |> 
  left_join(raw.pop |> select(popcode, eradication_status), by = "popcode")
```

```{r apparent eradication}
ggplot(pop.sel, aes(x = 1, fill = eradication_status)) +
  geom_bar(position = "stack") +
  facet_wrap(~ species, nrow = 2, ncol = 5) +
  labs(title = NULL,               # MODIFY
       x = NULL,
       y = "# populations") +
  scale_x_continuous(breaks = NULL) +
  guides(fill = guide_legend(title = NULL)) +
  theme_bw() +
  theme(legend.position = "bottom",
        strip.text = element_text(face = "italic"))
```

```{r save}
ggsave(filename = paste0("./data/output/KPI_", "pop_overview_", 
                         Sys.Date(), ".png"),
                         width = 15, height = 10, units= "cm")
```

# FUNCTIONS

```{r}
plot_this_absolute <- function(DATA){
  ggplot(DATA, aes(x = popcode)) +
    geom_segment(aes(xend = popcode,
                     y =    before_m2,
                     yend = after_m2.cap),
                 arrow = arrow(type = "closed", length = unit(0.15, "cm"))) +
    geom_hline(yintercept = 0.20,
               colour = "red",
               linetype = "dashed") +
    annotate("text",
             x = 1,
             y = 0.20 + 0.02,
             label = "target",
             hjust = 0,
             vjust = 0,
             size = 3.5,
             colour = "red") +
    labs(x = NULL,
         y = NULL,
         title = NULL) +
    theme_bw()
    # theme(
    #   legend.position = "bottom",
    #   plot.title = element_text(face = "italic"),
    #   axis.text.x = element_text(angle = 90, vjust = 0, hjust = 1))
  }
```

```{r}
plot_this_relative <- function(DATA){
  ggplot(DATA, aes(x = popcode)) +
    geom_segment(aes(xend = popcode,
                     y =    1,
                     yend = 1 - reduction.corr),
                 arrow = arrow(type = "closed", length = unit(0.15, "cm"))) +
    geom_hline(yintercept = 0.20,
               colour = "red",
               linetype = "dashed") +
    # annotate("text",
    #          x = 1,
    #          y = 0.20 + 0.02,
    #          label = "target",
    #          hjust = 0,
    #          vjust = 0,
    #          size = 3.5,
    #          colour = "red") +
    labs(x = NULL,
         y = NULL,
         title = NULL) +
    theme_bw()
    # theme(
    #   legend.position = "bottom",
    #   plot.title = element_text(face = "italic"),
    #   axis.text.x = element_text(angle = 90, vjust = 0, hjust = 1))
  }
```

```{r}
plot_this_absolute(pop.strip)
plot_this_absolute(pop.strip |> filter(popcode != "Ed3"))
plot_this_relative(pop.strip)
```

```{r save}
ggsave(filename = paste0("./data/output/KPI_", "pop_reduction_", 
                         Sys.Date(), ".png"),
                         width = 20, height = 5, units= "cm")
```