
Pro memorie :

meeting 2025-04-08  : https://drive.google.com/drive/folders/1J16lJqVCQ_Txaiix2DBOPZi1mdyYOQ4B
C2 log  : https://drive.google.com/drive/folders/1SwqbaRRxpQv4ByoCZskPje64NW5Es-vE?usp=sharing

Exact KPI formulation (at the end of the project):

End-of-project
KPI = Reduction of population size in managed sites of emerging plant species, from start of intervention (as baseline) to 2026.
Target = ≥ 80% decrease on baseline

5-years-after-project
KPI = A final assessment will be made at the end of the AfterLife project (2031) to check for complete population eradication.
Target = 100% decrease on baseline

Since a project baseline needs to be given as an absolute number, the choice is made to express these as number of m². This is taken as the sum of currently known populations under C2 management. For the populations currently expressed in number of plants (or other), a conversion factor will be applied (cover of a single plant).

# PREPARATION

```{r prep packages}
library(tidyverse)
library(readxl)
library(stringr)
library(sf)
library(leaflet)
library(mapview)
```

```{r prep memory}
rm(list = ls())
```

```{r prep handle coordinates}
dms_to_decimal <- function(dms) {
    deg_min_sec <- strsplit(dms, "[°'\",]")[[1]]
    deg <- as.numeric(deg_min_sec[1])
    min <- as.numeric(deg_min_sec[2])
    sec <- as.numeric(gsub(",", ".", deg_min_sec[3])) # vervang komma door punt
    hemisphere <- deg_min_sec[4]

    decimal <- deg + (min / 60) + (sec / 3600)
    return(decimal)
  }
```

# DATA

```{r read data}
file <- "./data/input/C2 RapidResponseActionsLog_asof26-08-28.xlsx" # most recent version

raw.pop <- read_excel(file, sheet = 2, skip = 1)
raw.act <- read_excel(file, sheet = 3, skip = 1)
```

```{r select}
pop.all <- raw.pop %>% 
  filter(group == "plants") |> 
  transmute(
    popcode,
    species = str_extract(species, "^[^ ]+"), # genus
    before_m2 = `INITIAL_m²`,
    before_other = INITIAL_other_quantity,
    before_other_unit = INITIAL_other_unit,
    after = REMAINING_size,
    after_unit = REMAINING_size_unit,
    eradication_status,
    visit = as.numeric(visit_base),
    lat = sapply(latitude, dms_to_decimal),
    lon = sapply(longitude, dms_to_decimal)
    )
```

# EXPLORE

## TAXA

```{r}
ggplot() +
  geom_bar(data = pop.all, aes(x = species)) +
  xlab(NULL) +
  ylab("# populations") +
  theme_bw()

# ggsave(filename = paste0("./data/output/KPI_overview_pops", Sys.Date(), ".jpg"),
#        width = 20, height = 10, units = "cm")
```

## MAP

```{r data}
rip <- st_read("./data/spatial/perimeter/Riparias_Official_StudyArea.geojson")

pop.plot <- pop.all |>
  st_as_sf(coords = c("lon", "lat"), crs = "+proj=longlat +datum=WGS84")
```

```{r simple plot}
ggplot() +
  geom_sf(data = rip) +
  geom_sf(data = pop.plot, aes(color = species)) +
  theme_bw() +
  theme(
    legend.position = "right",
    legend.title = element_blank()
    )

# ggsave(paste0("./data/output/KPI_pop-locs_plot_", Sys.Date(), ".jpg"))
```

```{r leaflet}
# map <-
leaflet() |> 
  setView(
    lng = 4.49,
    lat = 50.77,
    # lng = mean(st_coordinates(pop$geometry)[,1]),
    # lat = mean(st_coordinates(pop$geometry)[,2]),
    zoom = 10) |> 
  addTiles(group = "OSM") |> 
  addPolygons(data = rip,
              color = "blue",
              weight = 1.5,
              fillColor = "blue",
              fillOpacity = 0.1) |> 
  addCircleMarkers(data = pop.all,
                   radius = 5,
                   fillColor = "red",
                   opacity = 1,
                   fillOpacity = 1,
                   weight = 0.5,
                   label = ~popcode)
  
# mapshot(map, file = paste0("./data/output/KPI_pop-locs_leaflet_", Sys.Date(), ".jpg"),
#         remove_controls = c("zoomControl"))
```

# KPI

Now, quantitative data is needed.

```{r prep conversion factors}
La.plants.to.m2 <- 0.1 # Vibe # Note 1
Lp.plants.to.m2 <- 0.7 # Vibe
Lp.m3.to.m2     <- 6   # Vibe
Ch.m3.to.m2     <- 35  # Vibe
Ed.m3.to.m2     <- 4.6 # Vibe
Eg.plants.to.m2 <- 0.2 # Vibe
Pj.plants.to.m2 <- 1.8 # Vibe
Kp.plants.to.m2 <- 1.0 # Vibe
Zl.plants.to.m2 <- 0.1 # Vibe

# 1: was initially higher, but feedback from CR that plants were sampled late winter
```

```{r}
#  populations that were only managed from summer 2026 cannot be taken along
pop.kpi <- pop.all |> 
  filter(!visit == 1)

# Eg4, Eg5 : assume "<1" ~ 1
pop.kpi$before_m2[pop.kpi$before_m2 == "<1"] <- "1"

# Eg3 : data missing, let's assume 3m² (based on wnm.be and field visits)
pop.kpi$before_m2[pop.kpi$popcode == "Eg3"] <- "3"

# other
pop.kpi$after[pop.kpi$after == "n/a"] <- NA

pop.kpi$before_m2 <- as.numeric(pop.kpi$before_m2)
pop.kpi$after <- as.numeric(pop.kpi$after)
```

```{r}
pop.m2 <- pop.kpi %>%
  
  ## 'BEFORE' in M²
  # Lysichiton
  mutate(before_m2 = ifelse(species == "Lysichiton" &
                              is.na(before_m2) &
                              before_other_unit == "plants",
                            before_other * La.plants.to.m2,
                            before_m2)) |> 
  # Ludwigia
  mutate(before_m2 = ifelse(species == "Ludwigia" &
                              is.na(before_m2) &
                              before_other_unit == "plants",
                            before_other * Lp.plants.to.m2,
                            before_m2)) |> 
  mutate(before_m2 = ifelse(species == "Ludwigia" &
                              is.na(before_m2) &
                              before_other_unit == "m3",
                            before_other * Lp.m3.to.m2,
                            before_m2)) |> 
  # Crassula
  mutate(before_m2 = ifelse(species == "Crassula" &
                              is.na(before_m2) &
                              before_other_unit == "m3",
                            before_other * Ch.m3.to.m2,
                            before_m2)) |> 
  # Egeria
  mutate(before_m2 = ifelse(species == "Egeria" &
                              is.na(before_m2) &
                              before_other_unit == "m3",
                            before_other * Ed.m3.to.m2,
                            before_m2)) |> 

  ## 'AFTER' in M²
  mutate(after_m2 = NA, .before = after) |>
  # logic : if considered eradicated, then surface area = 0 m²
  mutate(after_m2 = ifelse(eradication_status == "ERADICATED",
                           0,
                           after_m2)) |>
  # if already expressed as m², then take that value
  mutate(after_m2 = ifelse(after_unit == "m2",
                           after,
                           after_m2)) |> 
  # convert from other units
  ## Lysichiton
  mutate(after_m2 = ifelse(species == "Lysichiton" &
                             is.na(after_m2) &
                             after_unit == "plants",
                            after * La.plants.to.m2,
                            after_m2)) |> 
  ## Ludwigia
  mutate(after_m2 = ifelse(species == "Ludwigia" &
                             is.na(after_m2) &
                             after_unit == "m3",
                            after * Lp.m3.to.m2,
                            after_m2)) |>
  mutate(after_m2 = ifelse(species == "Ludwigia" &
                             is.na(after_m2) &
                             after_unit == "plants",
                           after * Lp.plants.to.m2,
                           after_m2)) |> 
  ## Crassula
  mutate(after_m2 = ifelse(species == "Crassula" &
                             is.na(after_m2) &
                             after_unit == "m3",
                            after * Ch.m3.to.m2,
                            after_m2)) |> 
  ## Erythrante
  mutate(after_m2 = ifelse(species == "Erythranthe" &
                             is.na(after_m2) &
                             after_unit == "plants",
                            after * Eg.plants.to.m2,
                            after_m2)) |> 
  ## Petasites
  mutate(after_m2 = ifelse(species == "Petasites" &
                             is.na(after_m2) &
                             after_unit == "plants",
                            after * Pj.plants.to.m2,
                            after_m2)) |> 
  ## Koenigia
  mutate(after_m2 = ifelse(species == "Koenigia" &
                             is.na(after_m2) &
                             after_unit == "plants",
                            after * Kp.plants.to.m2,
                            after_m2)) |> 
  ## Zizania
  mutate(after_m2 = ifelse(species == "Zizania" &
                             is.na(after_m2) &
                             after_unit == "plants",
                           after * Zl.plants.to.m2,
                           after_m2)) |> 
  ## Egeria
  mutate(after_m2 = ifelse(species == "Egeria" &
                             is.na(after_m2) &
                             after_unit == "m3",
                           after * Ed.m3.to.m2,
                           after_m2))
```

```{r tdot}
missing <- pop.m2 |>
  filter(is.na(after_m2)) |>
  select(popcode) |>
  left_join(raw.pop |> select(popcode, site, region))
```

```{r strip it}
pop <- pop.m2 |> 
  # stump for filtering
  # filter() |> 
  select(popcode,
         species,
         before_m2,
         after_m2) |> 
  filter(!is.na(after_m2)) |> 
  mutate(
    reduction = 1 - (after_m2 / before_m2)
    ) |> 
  # accounting for the few populations that seem to have grown
  mutate(
    after_m2.cap = ifelse(after_m2 > before_m2,
                          before_m2,
                          after_m2), .after = after_m2) |>
  mutate(
    reduction.corr = 1 - (after_m2.cap / before_m2)
    )
```

# KPI

```{r kpi absolute}
# using the capped after_m2
kpi.1 <- 1 - sum(pop$after_m2.cap) / sum(pop$before_m2)
kpi.1
```

```{r relative}
kpi.2 <- mean(pop$reduction.corr)
kpi.2
```

# ERADICATION STATUS

```{r}
# Calculate total counts for each species and reorder
pop.bar <- pop.kpi %>%
  group_by(species) %>%
  mutate(total = n()) %>%
  ungroup()

pop.bar$species <- reorder(pop.bar$species,
                           pop.bar$total,
                           FUN = sum,
                           decreasing = TRUE)

pop.bar$eradication_status[pop.bar$eradication_status == "ERADICATED"] <- "absent"
pop.bar$eradication_status[pop.bar$eradication_status == "not"] <- "still present"

ggplot(pop.bar) +
  geom_bar(aes(x = species, fill = eradication_status), position = "stack") +
  labs(title = NULL,
       x = NULL,
       y = "# populations") +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        strip.text = element_text(face = "italic"))
```

```{r save}
ggsave(filename = paste0("./data/output/KPI_", "pop_overview_",
                         Sys.Date(), ".png"),
                         width = 20, height = 10, units= "cm")
```

# REDUCTION

```{r absolute reduction}
pop.arr <- pop |> 
  group_by(popcode) %>%
  mutate(order_var = before_m2) %>%
  ungroup()

pop.arr$popcode <- reorder(pop.arr$popcode, pop.arr$order_var,
                           FUN = max,
                           decreasing = TRUE)

ggplot(pop.arr, aes(x = popcode, color = species)) +
  geom_segment(aes(xend = popcode,
                   y =    before_m2,
                   yend = after_m2.cap),
               arrow = arrow(type = "closed", length = unit(0.15, "cm"))) +
  labs(x = NULL,
       y = NULL,
       title = NULL) +
  theme_bw() +
  scale_y_log10() +
  theme(
    axis.text.x = element_blank(),  # Remove x-axis labels
    # axis.ticks.x = element_blank(),  # Remove x-axis ticks
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
    ) +
  labs(
    x = "Population",
    y = "Population size (m²)"
    )

```

```{r save}
ggsave(filename = paste0("./data/output/KPI_", "pop_reduction_abs_",
                         Sys.Date(), ".png"),
       width = 20, height = 10, units = "cm")
```

```{r}
pop.arr2 <- pop |>
  group_by(popcode) %>%
  mutate(order_var = reduction.corr) %>%
  ungroup()

pop.arr2$popcode <- reorder(pop.arr2$popcode, pop.arr2$order_var,
                            FUN = max,
                            decreasing = TRUE)

ggplot(pop.arr2, aes(x = popcode, color = species)) +
  geom_segment(aes(xend = popcode,
                   y =    1,
                   yend = 1 - reduction.corr),
               arrow = arrow(type = "closed", length = unit(0.15, "cm"))) +
    geom_hline(yintercept = 0.20,
               colour = "red",
               linetype = "dashed") +
  labs(x = NULL,
       y = NULL,
       title = NULL) +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),  # Remove x-axis labels
    # axis.ticks.x = element_blank(),  # Remove x-axis ticks
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
    ) +
  labs(
    x = "Population",
    y = "Size (before = 1)"
    )
```

```{r save}
ggsave(filename = paste0("./data/output/KPI_", "pop_reduction_rel_",
                         Sys.Date(), ".png"),
       width = 20, height = 10, units = "cm")
```

# REGRESSIE

Doel: bepalen van interval rond gemiddelde reductie per soort.

Wegens hoge aanwezigheid van 0 en 1 : keuze voor betaregressie;

```{r}
library(betareg)
library(boot)
```

```{r}
pop.reg.data <- pop |> 
  filter(!species %in% c(
    "Lagarosiphon", # exclude those with 3 or less values
    "Koenigia",
    "Petasites",
    "Pontederia"
    )) |> 
  select(species, reduction.corr)

model <- betareg(data = pop.reg.data,
                 reduction.corr ~ species)
summary(model)
```

```{r}
fitted_values <- fitted(model)

mean_by_species <- aggregate(fitted_values,
                             by = list(species = pop.reg.data$species),
                             FUN = mean)
print(mean_by_species)
```

# ALTERNATIEF : GEMIDDELDE EN ST.ERR

```{r}
pop.summ.raw <- pop |> 
  group_by(species) |> 
  summarize(
    n = n(),
    red.mean = mean(reduction.corr),
    red.std  = sd(reduction.corr)
    ) |> 
  mutate(
    red.se = red.std/sqrt(n)
    )

pop.summ <- pop.summ.raw |> 
  select(species, n, red.mean, red.se) |> 
  filter(n > 3) |> 
  arrange(desc(red.mean))

pop.summ$species <- fct_reorder(pop.summ$species,
                                pop.summ$red.mean,
                                .desc = TRUE)

```

```{r}
ggplot(data = pop.summ) +
  geom_pointrange(
    aes(
      x = species,
      y = red.mean,
      ymin = red.mean - red.se, # estimate-se,
      ymax = red.mean + red.se, # estimate+se
      )
    ) +
  theme_bw() +
  labs(
    x = NULL,
    y = "Size reduction"
    )
```

```{r save}
ggsave(filename = paste0("./data/output/KPI_", "pop_reduction_mean_",
                         Sys.Date(), ".png"),
       width = 10, height = 8, units = "cm")
```