
Pro memorie :

meeting 2025-04-08  : https://drive.google.com/drive/folders/1J16lJqVCQ_Txaiix2DBOPZi1mdyYOQ4B
C2 log  : https://drive.google.com/drive/folders/1SwqbaRRxpQv4ByoCZskPje64NW5Es-vE?usp=sharing

Exact KPI formulation (at the end of the project):

End-of-project
KPI = Reduction of population size in managed sites of emerging plant species, from start of intervention (as baseline) to 2026.
Target = ≥ 80% decrease on baseline

5-years-after-project
KPI = A final assessment will be made at the end of the AfterLife project (2031) to check for complete population eradication.
Target = 100% decrease on baseline

Since a project baseline needs to be given as an absolute number, the choice is made to express these as number of m². This is taken as the sum of currently known populations under C2 management. For the populations currently expressed in number of plants (or other), a conversion factor will be applied (cover of a single plant).

# PREPARATION

```{r prep packages}
library(tidyverse)
library(readxl)
library(stringr)
```

```{r prep memory}
rm(list = ls())
```

# DATA

```{r read data}
file <- "./data/input/C2 RapidResponseActionsLog_asof26-08-04.xlsx" # most recent version

raw.pop <- read_excel(file, sheet = 2, skip = 1)
raw.act <- read_excel(file, sheet = 3, skip = 1)
```

```{r select}
pop <- raw.pop %>% 
  filter(group == "plants") |> 
  transmute(
    popcode,
    species = str_extract(species, "^[^ ]+"), # genus
    # latitude,
    # longitude,
    # region,
    before_m2 = `INITIAL_infested_m²`,
    before_other = INITIAL_size_quantity,
    before_other_unit = INITIAL_size_unit,
    after = REMAINING_size,
    after_unit = REMAINING_size_unit,
    eradication_status
    )
```

```{r}
spp <- unique(pop$species)
```

```{r fix and complete}

# Eg4, Eg5 : assume "<1" ~ 1
pop$before_m2[pop$before_m2 == "<1"] <- "1"

# Eg3 : data missing, let's assume 3m² (based on wnm.be and field visits)
pop$before_m2[pop$popcode == "Eg3"] <- "3"

# other
pop$after[pop$after == "n/a"] <- NA

pop$before_m2 <- as.numeric(pop$before_m2)
pop$after <- as.numeric(pop$after)

```

# CONVERT ALL TO M²

```{r}
La.plants.to.m2 <- 1.2 # Vibe
Lg.plants.to.m2 <- 0.7 # Vibe
Lg.m3.to.m2     <- 6   # Vibe
Ch.m3.to.m2     <- 35  # Vibe
Ed.m3.to.m2     <- 4.6 # Vibe
Eg.plants.to.m2 <- 0.2 # Vibe
Pj.plants.to.m2 <- 1.8 # Vibe
Kp.plants.to.m2 <- 1.0 # Vibe
```

## 'BEFORE' in M²

```{r}
pop.m2 <- pop %>%
  # Lysichiton
  mutate(before_m2 = ifelse(species == "Lysichiton" &
                              is.na(before_m2) &
                              before_other_unit == "plants",
                            before_other * La.plants.to.m2,
                            before_m2)) |> 
  # Ludwigia
  mutate(before_m2 = ifelse(species == "Ludwigia" &
                              is.na(before_m2) &
                              before_other_unit == "plants",
                            before_other * Lg.plants.to.m2,
                            before_m2)) |> 
  mutate(before_m2 = ifelse(species == "Ludwigia" &
                              is.na(before_m2) &
                              before_other_unit == "m3",
                            before_other * Lg.m3.to.m2,
                            before_m2)) |> 
  # Crassula
  mutate(before_m2 = ifelse(species == "Crassula" &
                              is.na(before_m2) &
                              before_other_unit == "m3",
                            before_other * Ch.m3.to.m2,
                            before_m2)) |> 
  # Egeria
  mutate(before_m2 = ifelse(species == "Egeria" &
                              is.na(before_m2) &
                              before_other_unit == "m3",
                            before_other * Ed.m3.to.m2,
                            before_m2))
```

## 'AFTER' in M²

```{r}
pop.m2 <- pop.m2 |> 
  mutate(after_m2 = NA, .before = after) |>
  # logic : if considered eradicated, then surface area = 0 m²
  mutate(after_m2 = ifelse(eradication_status == "ERADICATED",
                           0,
                           after_m2)) |>
  # Lysichiton
  mutate(after_m2 = ifelse(species == "Lysichiton" &
                             is.na(after_m2) &
                             after_unit == "plants",
                            after * La.plants.to.m2,
                            after_m2)) |> 
  # Ludwigia
  mutate(after_m2 = ifelse(species == "Ludwigia" &
                             is.na(after_m2) &
                             after_unit == "m3",
                            after * Lg.m3.to.m2,
                            after_m2)) |>
  # Crassula
  mutate(after_m2 = ifelse(species == "Crassula" &
                             is.na(after_m2) &
                             after_unit == "m3",
                            after * Ch.m3.to.m2,
                            after_m2)) |> 
  # Erythrante
  mutate(after_m2 = ifelse(species == "Erythranthe" &
                             is.na(after_m2) &
                             after_unit == "m2",
                           after,
                           after_m2)) |> 
  mutate(after_m2 = ifelse(species == "Erythranthe" &
                             is.na(after_m2) &
                             after_unit == "plants",
                            after * Eg.plants.to.m2,
                            after_m2)) |> 
  # Petasites
  mutate(after_m2 = ifelse(species == "Petasites" &
                             is.na(after_m2) &
                             after_unit == "plants",
                            after * Pj.plants.to.m2,
                            after_m2)) |> 
  # Koenigia
  mutate(after_m2 = ifelse(species == "Koenigia" &
                             is.na(after_m2) &
                             after_unit == "plants",
                            after * Kp.plants.to.m2,
                            after_m2))
```

```{r tdot}
missing <- pop.m2 |> 
  filter(is.na(after_m2)) |> 
  select(popcode) |> 
  write.csv2("./data/missing_asof26-08-04_td.csv")
```

```{r strip it}
pop.strip <- pop.m2 |> 
  select(popcode,
         species,
         before_m2,
         after_m2) |> 
  filter(!is.na(after_m2)) |> 
  mutate(
    reduction = 1 - (after_m2 / before_m2)
    ) |> 
  # accounting for the few populations that seem to have grown
  mutate(
    after_m2.cap = ifelse(after_m2 > before_m2,
                          before_m2,
                          after_m2), .after = after_m2) |>
  mutate(
    reduction.corr = ifelse(reduction < 0,
                            0,
                            reduction)
    )
```

```{r naive method}
# using the capped after_m2
kpi.1 <- 1 - sum(pop.strip$after_m2.cap) / sum(pop.strip$before_m2)
kpi.1
```

```{r per-population method}
# 2do
```

****************************************************** OLD - TO DELETE

# QUICK OVERVIEW

```{r apparent eradication}
ggplot(pop, aes(x = 1, fill = eradication_status)) +
  geom_bar(position = "stack") +
  facet_wrap(~ genus, nrow = 2, ncol = 5) +
  labs(title = "status as of 04/03/2026",               # MODIFY
       x = NULL,
       y = "# populations") +
  scale_x_continuous(breaks = NULL) +
  guides(fill = guide_legend(title = NULL)) +
  theme_bw() +
  theme(legend.position = "bottom",
        strip.text = element_text(face = "italic"))
```

```{r save}
ggsave(filename = paste0("./data/output/KPI_", "pop_overview_", 
                         Sys.Date(), ".png"),
                         width = 15, height = 15, units= "cm")
```

# FUNCTIONS

```{r}
plot_this_species <- function(DATA){
  ggplot(DATA, aes(x = popcode)) +
    geom_segment(aes(xend = popcode,
                     y =    size.rel.bas,
                     yend = size.rel.eop,
                     colour = status),
                 arrow = arrow(type = "closed", length = unit(0.15, "cm"))) +
    geom_hline(yintercept = 0.20,
               colour = "red",
               linetype = "dashed") +
    annotate("text",
             x = 1,
             y = 0.20 + 0.02,
             label = "target",
             hjust = 0,
             vjust = 0,
             size = 3.5,
             colour = "red") +
    scale_colour_manual(values = c(confirmed = "black",
                                   confirmation_needed = "blue"),
                        name = NULL) +
    labs(x = NULL,
         y = NULL,
         title = first(DATA$species)) +
    theme_bw() +
    theme(
      legend.position = "bottom",
      plot.title = element_text(face = "italic"),
      axis.text.x = element_text(angle = 90, vjust = 0, hjust = 1))
  }
```

# PLANT-BASED

= voor waar de eenheid het AANTAL PLANTEN is

```{r}
pop.short.1 <- pop %>%
  transmute(popcode,
            species,
            size.unit = pop_size_unit,
            size.bas = pop_size_quantity, # bas = baseline
            eradication_status,
            status,
            size.eop = NA                 # eop = end-of-project
            )
```

## LYS

```{r}
lys <- pop.short.1 %>% 
  filter(species == "Lysichiton americanus") %>%  
  
  # considered eradicated
  mutate(size.eop = ifelse(
    eradication_status == "apparent_eradication", 0, size.eop)) %>%
  
  # others, in follow-up phase (ad hoc solutions)
  mutate(size.eop = case_when(
    popcode == "La2" ~ 15 - 15,          # all plants
    popcode == "La5" ~ 200 - 150 - 18,   # removed 2024 + '25
    popcode == "La7" ~ 70 - 70,          # all plants
    popcode == "La8" ~ 30 - 30,          # all plants
    popcode == "La12" ~ 10 - 10,         # all plants
    popcode == "La14" ~ 7 - 7,           # "
    popcode == "La15" ~ 30 - 30,         # "
    popcode == "La17" ~ 10 - 10,         # "
    popcode == "La18" ~ 30 - 25,         # 2025
    popcode == "La21" ~ 60 - 60,         # all plants
    popcode == "La24" ~ 50 - 50,         # all plants
    popcode == "La25" ~ 2 - 1,           # 2025
    popcode == "La26" ~ 50 - 50,         # all plants
    TRUE ~ size.eop
  )) %>% 
  
  mutate(size.rel.bas = size.bas/size.bas,
         size.rel.eop = size.eop/size.bas) %>% 
  arrange(size.rel.eop)
```

```{r}
plot_this_species(lys)
```

```{r save}
ggsave(filename = paste0("./data/output/KPI_", "Lysichiton_", 
                         Sys.Date(), ".png"),
                         width = 15, height = 8, units= "cm")
```

# AREA-BASED

= voor waar de eenheid het AANTAL PLANTEN is

```{r}
pop.short.2 <- pop %>%
  transmute(popcode,
            species,
            size.bas = `infested_area_m²`, # bas = baseline
            eradication_status,
            status,
            size.eop = NA                 # eop = end-of-project
            )
```

## ZIZ

```{r}
ziz <- pop.short.2 %>% 
  filter(species == "Zizania latifolia") %>%  
  
  # considered eradicated
  mutate(size.eop = ifelse(
    eradication_status == "apparent_eradication", 0, size.eop)) %>%
  
  # others, in follow-up phase (ad hoc solutions)
  mutate(size.eop = case_when(
    popcode == "Zl1" ~ 843 - 0.1*843, # dummy
    popcode == "Zl2" ~ 196.5 - 196.5, # 2024
    popcode == "Zl4" ~ 1570.9 - 0.1*1570.9, # dummy
    # popcode == "Zl5" ~ NA_character_,
    popcode == "Zl6" ~ 1300 - 675, # 2025
    popcode == "Zl7" ~ 270 - 0.1*270, # dummy
    popcode == "Zl8" ~ 15 - 15, # 2025
    TRUE ~ size.eop
  )) %>% 
  
  mutate(size.rel.bas = size.bas/size.bas,
         size.rel.eop = size.eop/size.bas) %>% 
  arrange(size.rel.eop)
```

```{r}
plot_this_species(ziz)
```

```{r save}
ggsave(filename = paste0("./data/output/KPI_", "Zizania", 
                         Sys.Date(), ".png"),
                         width = 15, height = 8, units= "cm")
```

# PON

```{r}
pon <- pop.short.2 %>% 
  filter(species == "Pontederia cordata") %>%  
  
  # considered eradicated
  mutate(size.eop = ifelse(
    eradication_status == "apparent_eradication", 0, size.eop)) %>%
  
  # others, in follow-up phase (ad hoc solutions)
  mutate(size.eop = case_when(
    popcode == "Pco1" ~ 3 - 3, # all plants
    popcode == "Pco2" ~ 14.7 - 4, # dummy
    TRUE ~ size.eop
  )) %>% 
  
  mutate(size.rel.bas = size.bas/size.bas,
         size.rel.eop = size.eop/size.bas) %>% 
  arrange(size.rel.eop)
```

```{r}
plot_this_species(pon)
```

```{r save}
ggsave(filename = paste0("./data/output/KPI_", "Pontederia", 
                         Sys.Date(), ".png"),
                         width = 15, height = 8, units= "cm")
```

# LUD

# CRA

```{r}
cra <- pop.short.2 %>% 
  filter(species == "Crassula helmsii") %>%  
  
  # considered eradicated
  mutate(size.eop = ifelse(
    eradication_status == "apparent_eradication", 0, size.eop)) %>%
  
  # others, in follow-up phase (ad hoc solutions)
  mutate(size.eop = case_when(
    popcode == "Ch1" ~ 200 - 0.1*200, # dummy
    popcode == "Ch2" ~ 10 - 10, # all plants
    popcode == "Ch3" ~ 200 - 200, # all plants
    popcode == "Ch5" ~ 3 - 0.1*3, # dummy
    popcode == "Ch6" ~ 40 - 0.1*40, # dummy
    TRUE ~ size.eop
  )) %>% 
  
  mutate(size.rel.bas = size.bas/size.bas,
         size.rel.eop = size.eop/size.bas) %>% 
  arrange(size.rel.eop)
```

```{r}
plot_this_species(cra)
```

```{r save}
ggsave(filename = paste0("./data/output/KPI_", "Crassula", 
                         Sys.Date(), ".png"),
                         width = 15, height = 8, units= "cm")
```

# LAG

```{r}
lag <- pop.short.2 %>% 
  filter(species == "Lagarosiphon major") %>%  
  
  # considered eradicated
  mutate(size.eop = ifelse(
    eradication_status == "apparent_eradication", 0, size.eop)) %>%
  
  # others, in follow-up phase (ad hoc solutions)
  # mutate(size.eop = case_when(
  #       TRUE ~ size.eop
  # )) %>% 
  
  mutate(size.rel.bas = size.bas/size.bas,
         size.rel.eop = size.eop/size.bas) %>% 
  arrange(size.rel.eop)
```

```{r}
plot_this_species(lag)
```

```{r save}
ggsave(filename = paste0("./data/output/KPI_", "Lagarosiphon", 
                         Sys.Date(), ".png"),
                         width = 15, height = 8, units= "cm")
```

# EGE

```{r}
ege <- pop.short.2 %>% 
  filter(species == "Egeria densa") %>%  
  
  # considered eradicated
  mutate(size.eop = ifelse(
    eradication_status == "apparent_eradication", 0, size.eop)) %>%
  
  # others, in follow-up phase (ad hoc solutions)
  mutate(size.eop = case_when(
    popcode == "Ed2" ~ 80 - 0.9*80, # dummy
    popcode == "Ed3" ~ 2890 - 0.9*2890, # all plants
    TRUE ~ size.eop
    )) %>%
  
  mutate(size.rel.bas = size.bas/size.bas,
         size.rel.eop = size.eop/size.bas) %>% 
  arrange(size.rel.eop)
```

```{r}
plot_this_species(ege)
```

```{r save}
ggsave(filename = paste0("./data/output/KPI_", "Egeria", 
                         Sys.Date(), ".png"),
                         width = 15, height = 8, units= "cm")
```

# ERY

```{r}
ery <- pop.short.2 %>% 
  filter(species == "Erythranthe guttata") %>%  
  
  # considered eradicated
  mutate(size.eop = ifelse(
    eradication_status == "apparent_eradication", 0, size.eop)) %>%
  
  # others, in follow-up phase (ad hoc solutions)
  mutate(size.eop = case_when(
    popcode == "Eg1" ~ 10 - 0.95*10, # comment
    popcode == "Eg2" ~ 200 - 0.95*200, # all plants
    TRUE ~ size.eop
    )) %>%
  
  mutate(size.rel.bas = size.bas/size.bas,
         size.rel.eop = size.eop/size.bas) %>% 
  arrange(size.rel.eop)
```

```{r}
plot_this_species(ery)
```

```{r save}
ggsave(filename = paste0("./data/output/KPI_", "Erythranthe", 
                         Sys.Date(), ".png"),
                         width = 15, height = 8, units= "cm")
```

# PET

```{r}
pet <- pop.short.2 %>% 
  filter(species == "Petasites japonicus") %>%  
  
  # considered eradicated
  mutate(size.eop = ifelse(
    eradication_status == "apparent_eradication", 0, size.eop)) %>%
  
  # others, in follow-up phase (ad hoc solutions)
  mutate(size.eop = case_when(
    popcode == "Pj1" ~ 85 - 85, # all plants
    TRUE ~ size.eop
    )) %>%
  
  mutate(size.rel.bas = size.bas/size.bas,
         size.rel.eop = size.eop/size.bas) %>% 
  arrange(size.rel.eop)
```

```{r}
plot_this_species(pet)
```

```{r save}
ggsave(filename = paste0("./data/output/KPI_", "Petasites", 
                         Sys.Date(), ".png"),
                         width = 15, height = 8, units= "cm")
```

# KOE

## GENERAL STATICS (all species)

```{r}
all <- rbind(
  ziz %>% select(popcode, species, size.rel.bas, size.rel.eop),
  lys %>% select(popcode, species, size.rel.bas, size.rel.eop),
  pon %>% select(popcode, species, size.rel.bas, size.rel.eop),
  # lud, # nog te doen?
  cra %>% select(popcode, species, size.rel.bas, size.rel.eop),
  lag %>% select(popcode, species, size.rel.bas, size.rel.eop),
  ege %>% select(popcode, species, size.rel.bas, size.rel.eop),
  ery %>% select(popcode, species, size.rel.bas, size.rel.eop),
  pet %>% select(popcode, species, size.rel.bas, size.rel.eop)
  # koe # nog te doen?
  ) %>% 
  mutate(decline = size.rel.bas - size.rel.eop)
```

```{r}
ggplot(data = all) +
  geom_point(aes(x = popcode, y = size.rel.eop)) +
  facet_wrap(~ species, nrow = 2, ncol = 4) +
  theme_bw()
```

```{r save}
ggsave(filename = paste0("./data/output/KPI_", "all_", 
                         Sys.Date(), ".png"),
                         width = 15, height = 8, units= "cm")
```

# the acutual kpi

```{r kpi}
mean(all$decline, na.rm = T)
```

